Hydras \& Co.: Formalized mathematics in Coq\\ for inspiration and entertainment

### 2022-04-27 doc/JFLA2022/Slides/Slides.tex
work on JFLA slides

\documentclass[10pt]{beamer}
\usetheme{Montpellier}

@@

    Using the lemmas generated by \emph{coq-equations}, we prove minoration lemmas or exact formulas for given $\alpha$\,s:

    \begin{itemize}
    \item \input{../../movies/snippets/Hprime/HprimeOmegaCube3c}
      \input{../../movies/snippets/Hprime/HprimeOmegaCube3e}
     \item   \input{../../movies/snippets/Hprime/HprimeHexp2}
    \item
      \input{../../movies/snippets/Hprime/HprimePhi0OmegaClosed}

      \end{itemize}

@@

  \begin{block}{}
    Using combinatorial lemmas by  K. \& S., we prove (by mutual  transfinite induction on \textcolor{mathcolor}{$\alpha$}):
    \begin{quote}
      \begin{itemize}
      \item \textcolor{mathcolor}{$H'_\alpha$} is strictly monotonous
      \item If $\alpha>0$ then \textcolor{mathcolor}{$ i < H'_\alpha(i)$} for all $i$
        \item $H'_\alpha(i) < H'_{\alpha+1}(i) \; (i \not=0)$
      \item   If \textcolor{mathcolor}{$\alpha \rplus{n,n,\dots,n} \beta$} then \textcolor{mathcolor}{$H'_\beta(n)\leq H'_\alpha(n)$} 
      \end{itemize}
    \end{quote}

  \end{block}
  \begin{block}{Theorem}
    The function which computes the length of hydra battles
    (according to the initial index) is not primitive recursive
    (for $\alpha\geq \omega^\omega$).
    \vspace{4pt}

@@

  \begin{scriptsize}
    \inputsnippets{./SchemePrimRecInda}
  \end{scriptsize}
\end{frame}

@@

       \input{../../movies/snippets/F_omega/lemmas}

  \end{block}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
  \frametitle{Hydras \& Gaia}
  \begin{block}{From \gaia to \Hydras}
    {\footnotesize
    \inputsnippets{./multA}}
  \end{block}
\end{frame}

\begin{frame}
  \begin{block}{From \Hydras to \gaia}
    {\footnotesize
      \inputsnippets{./canonDef}
      \inputsnippets{./gcanonLimitOf}}
  \end{block}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Project Organization}
\begin{frame}
  \frametitle{Project  Organization}
  \begin{itemize}
  \item Integration with \community
  \item Documentation with \alectr
  \item Package dependencies
  \item Continuous integration
  \end{itemize}
\end{frame}

@@

\begin{frame}[fragile]
\begin{alltt}
{\color{lightblue}{(* begin snippet Ex42a:: no-out *)}}
Example Ex42: omega + 42 + omega^2 = omega^2. 
{\color{lightblue}{(* end snippet Ex42a *)}}
Proof.
  {\color{lightblue}{(* begin snippet Ex42b *)}}
  assert (HAP:= AP_phi0 2). {\color{lightblue}{(* .no-out *)}}
  elim  HAP; intros _ H0; apply H0; clear H0. 
  {\color{lightblue}{(* end snippet Ex42b *)}}
  {\color{lightblue}{(* begin snippet Ex42d *)}}
  Check AP_plus_closed. {\color{lightblue}{(* .unfold .no-goals *)}}
  {\color{lightblue}{(* end snippet Ex42d *)}}
   {\color{lightblue}{(* begin snippet Ex42c *)}}
  assert (Hlt: omega < omega^2) by 
      (rewrite omega_eqn; apply phi0_mono, finite_mono;
        auto with arith).
 {\color{lightblue}{ (* end  snippet Ex42c *)}}
\end{alltt}

@@

\begin{frame}
  \begin{small}
    \inputsnippets{
      ../Schutte_Ex42c, ../Schutte_Ex42d, ../Schutte_Ex42e}
  \end{small}

>>>>

\documentclass[10pt, fleqn]{beamer}
\usetheme{Montpellier}

@@

      {\color{mathcolor}

    \begin{align*}
      H'_{\omega^3}(k) \geq\, & \textit{hyper\_exp2}(k+1) \\
      H'_{\omega^\omega}(k)=\,& \textbf{let}\,F\;f\;i := f^{(i+1)}(i)\,
                      \textbf{in}\, F^{(k)}(\lambda k.\,2k+1)\,k\\
    \end{align*}
       }%mathcolor
  \end{block}

\begin{block}{Lemmas (after K.\& S.)}
   \begin{itemize} 
    \item \textcolor{mathcolor}{$H'_\alpha$} is strictly monotonous

     \item  If $\alpha>0$ then \textcolor{mathcolor}{$ i < H'_\alpha(i)$} for all $i$

    \item   $H'_\alpha(i) < H'_{\alpha+1}(i) \; (i \not=0)$

       \item     If \textcolor{mathcolor}{$\alpha \rplus{n,n,\dots,n} \beta$} then \textcolor{mathcolor}{$H'_\beta(n)\leq H'_\alpha(n)$}

       \end{itemize}
       (Proof by mutual transfinite induction)
 \end{block}

@@

  \begin{block}{Theorem}
    The function which computes the length of hydra battles
    (according to the initial index) is not primitive recursive
    (for $\alpha\geq \omega^\omega$).
    \vspace{4pt}

@@

\begin{frame}
  \frametitle{Proof sketch}
  \begin{scriptsize}
    \inputsnippets{./SchemePrimRecInda}
  \end{scriptsize}
\end{frame}

@@

Comparison of the length of hydra-battles (w.r.t to the initial replication factor) with the diagonalized Ackermann function proves the theorem.
  \end{block}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}
  \frametitle{Compatibility with Gaia (José Grimm)}
  \begin{block}{From \Hydras to \gaia}
    {\footnotesize
      \inputsnippets{./canonDef}
      \inputsnippets{./gcanonLimitOf}}
  \end{block}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Project Organization}
\begin{frame}
  \frametitle{Project  Organization}
  \begin{itemize}
  \item Integration with \community
  \item Documentation with \alectr
  \item Package dependencies
  \item Continuous integration
  \end{itemize}
\end{frame}

@@

\begin{frame}[fragile]
\begin{alltt}
{\color{lightblue}{(* begin snippet Ex42a:: no-out *)}}
Example Ex42: omega + 42 + omega^2 = omega^2. 
{\color{lightblue}{(* end snippet Ex42a *)}}
Proof.
  {\color{lightblue}{(* begin snippet Ex42b *)}}
  assert (HAP:= AP_phi0 2). {\color{lightblue}{(* .no-out *)}}
  elim  HAP; intros _ H0; apply H0; clear H0. 
  {\color{lightblue}{(* end snippet Ex42b *)}}
 \end{alltt}

@@
